---
title: "IF_analysis"
output:
  html_document:
    df_print: paged
date: "2024-07-26"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Adding packages

```{r}
library(readxl)
library(ggplot2)
library(tidyverse)
library(lubridate)
library(RColorBrewer)
```


## Statistical analyses

```{r}
#various packages related to mixed modelling
library(lme4)
library(nlme)
library(lattice)
library(effects)
library(emmeans)
library(car)
```


## Amy setwd and upload data

```{r}
setwd("C:/Users/amyh/Documents/R_data/2024_internships")
taxa <- read.csv("Troy_data2.csv")
water_quality <- read.csv("Troy_data.csv")


```


## Datasets, Palettes

Reading in the data sets, and set up colorblind-friendly color palette(s)

```{r}
taxa <- read_excel("~/Desktop/Iron Floc UA Project/iron_floc_impact_on_macroinvertebrate_biodiversity_taxa.xlsx")

water_quality <- read_excel("~/Desktop/Iron Floc UA Project/iron_floc_impact_on_macroinvertebrate_biodiversity_water_quality.xlsx", 
    sheet = "data")

cbPalette <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
```


## Data Formats

If you are getting errors and the syntax/et are correct, check what type of data the colum is, and run an as.DATATYPE if it's the wrong kind of data for what you're trying to do.

```{r}
str(taxa)
str(water_quality)
```


## Creating the Environment

Create the all_sites dataframe to summarize the site_name, analogue, number of species, number of individuals, biodiversity/species richness percent, and mean iron floc coverage percent.

```{r}
prcnt_cvr <- water_quality %>%
  group_by(site_name, condition) %>%
  summarize(mean_cover=mean(as.numeric(percent_cover), na.rm = TRUE), flow=mean(as.numeric(flow_rate)), ferrous_iron=mean(as.numeric(ferrous_iron)))

##this created all_sites, if you need to summarize the total number of unique/total number of entries
all_sites <- taxa %>% group_by(site_name, analogue) %>%
  ##select(site_name, taxa) %>% not necesary, try if there are errors
  summarize(count_taxa=n_distinct(taxa), total=n()) %>%
  mutate(biodiversity=count_taxa/total)
##adds the biodiversity values

sum_sites <- merge(prcnt_cvr, all_sites)
```




## Question 1

Does stream restoration increase the presence of iron flocculate by creating low-flow, acidic, and high iron water conditions compared to unrestored streams?
-%cover
-flow_rate
-ferrous-iron
x-condition

AMY: Additional Data manipulation/exploration
1. Percent Cover in restored/unrestored 
```{r}
#only using "at" plot locations
water_quality1 <- water_quality %>%
  filter(plot_location == "at")

#water_quality1 <- as.data.frame(water_quality1)

#percent cover from non-numeric to numeric
water_quality1$percent_cover <- as.numeric(water_quality1$percent_cover)

#BarGraphStats for calculating means
barGraphStats <- function(data, variable, byFactorNames) {
  count <- length(byFactorNames)
  N <- aggregate(data[[variable]], data[byFactorNames], FUN=length)
  names(N)[1:count] <- byFactorNames
  names(N) <- sub("^x$", "N", names(N))
  mean <- aggregate(data[[variable]], data[byFactorNames], FUN=mean)
  names(mean)[1:count] <- byFactorNames
  names(mean) <- sub("^x$", "mean", names(mean))
  sd <- aggregate(data[[variable]], data[byFactorNames], FUN=sd)
  names(sd)[1:count] <- byFactorNames
  names(sd) <- sub("^x$", "sd", names(sd))
  preSummaryStats <- merge(N, mean, by=byFactorNames)
  finalSummaryStats <- merge(preSummaryStats, sd, by=byFactorNames)
  finalSummaryStats$se <- finalSummaryStats$sd / sqrt(finalSummaryStats$N)
  return(finalSummaryStats)
}

#ggplot(g1 <- ggplot(data = barGraphStats(data = water_quality1, variable = "percent_cover", byFactorNames = c("site_name")), aes(x=site_name, y=mean)) +
    #geom_bar(stat='identity', position=position_dodge(), width=0.8) +
    #geom_errorbar(aes(ymin=mean-se, ymax=mean+se), width = 0.2, position = position_dodge(0.9)) +
    #annotate("text", x = 1, y = 2.7 , label = "a", size = 8)+
    #annotate("text", x = 2, y = 2.7 , label = "a", size = 8)+
    #annotate("text", x = 3, y = 1.75, label = "b", size = 8) +
   #theme(legend.position = "none") + xlab("Restoration") + ylab("Percent Cover"))
```


Preliminary Stats
```{r}
#Q1

#### Testing for normality ####
plot(water_quality1$percent_cover)
hist(water_quality1$percent_cover)
shapiro.test(water_quality1$percent_cover)
#data not normal

#log transform
water_quality1$percent_cover_log <- log(water_quality1$percent_cover +1)
shapiro.test(water_quality1$percent_cover_log)
plot(water_quality1$percent_cover)
hist(water_quality1$percent_cover)
#still not normal

#logit transform
water_quality1$percent_cover_logit <- logit(water_quality1$percent_cover)
shapiro.test(water_quality1$percent_cover_logit)
plot(water_quality1$percent_cover_logit)
hist(water_quality1$percent_cover_logit)
#still not normal, going to use OG data set for time's sake


mod1 <- lm(percent_cover ~ flow_rate*condition*ferrous_iron, data = water_quality1)

summary(mod1)
anova(mod1)

plot(mod1)

plot(water_quality1$percent_cover ~ water_quality1$flow_rate)

```

Percent Cover vs. Flow rate
```{r}
ggplot(water_quality1, aes(x=factor(flow_rate), y=percent_cover, fill=factor(condition)))+
  geom_point()+
  scale_color_manual(c("pink", "green"))+
  labs(x="Flow Rate (m^3/s)", y="Percent Cover of Iron Flocculant", fill="Restored/Unrestored")
```

Q1 Graphs
Percent cover, dissolved iron, and flow rate against site name filled by condition

```{r}
ggplot(sum_sites, aes(x=factor(site_name), y=mean_cover, fill=factor(condition)))+
  geom_bar(stat="identity")+
  #geom_text(aes(label=mean_cover), vjust=-1)+
  scale_x_discrete(limits = c("j1", "j3", "cc", "ccr", "ar", "hb"))+
  scale_y_continuous(labels=scales::percent, limits=c(0, 1))+
  labs(x="Site Name", y="Percent Cover", fill="Condition")
 
ggplot(sum_sites, aes(x=factor(site_name), y=ferrous_iron, fill=factor(condition)))+
  geom_bar(stat="identity")+
  #geom_text(aes(label=ferrous_iron), vjust=-1)+
  scale_x_discrete(limits = c("j1", "j3", "cc", "ccr", "ar", "hb"))+
  scale_y_continuous(limits=c(0, 15))+
  labs(x="Site Name", y="Dissolved Ferrous Iron", fill="Condition")

ggplot(sum_sites, aes(x=factor(site_name), y=flow, fill=factor(condition)))+
  geom_bar(stat="identity")+
  #geom_text(aes(label=mean_cover), vjust=-1)+
  scale_x_discrete(limits = c("j1", "j3", "cc", "ccr", "ar", "hb"))+
  #scale_y_continuous(labels=scales::percent, limits=c(0, 1))+
  labs(x="Site Name", y="Flow Rate (m^3/s)", fill="Condition")

ggplot(sum_sites, aes(x=flow, y=mean_cover))+
  geom_point(aes(color=condition, shape=condition), size=4)+
  geom_smooth(method="lm")+
  labs(x="Flow Rate", y="Iron Floc Percent Cover", title="Iron Floc as a function of Flow Rate", caption="r-squared=0.57 p=0.04")+
  theme(legend.direction="horizontal", legend.position="bottom")

ggplot(sum_sites, aes(x=ferrous_iron, y=mean_cover))+
  geom_point(aes(color=condition, shape=condition), size=4)+
  geom_smooth(method="lm")+
  labs(x="Dissolved Ferrous Iron (ppm)", y="Iron Floc Percent Cover", title="Iron Floc as a function of Dissolved Iron")+
  theme(legend.direction="horizontal", legend.position="bottom")
```


## Question 2

How does iron flocculate affect the richness and abundance of macroinvertebrates?

Plot number of species and number of individuals by percent cover and ferrous iron. Four graphs total. A fifth graph can show the biodiversity number against percent cover/ferrous iron.
---Add geom_smooth, but the smooth doesn't show due to scale manipulation. How best to show statistical significance?

```{r}
ggplot(sum_sites, aes(x=total, y=mean_cover))+
  geom_point(aes(color=condition, shape=condition), size=3)+
  geom_smooth(method="lm")+
  scale_y_continuous(labels=scales::percent, limits=c(0, 1))+
  labs(title="Abundance as a function of Iron Floc Coverage", x="Abundance", y="Fe Floc Percent Cover", color="Condition", shape="Condition")+
    theme(legend.direction="horizontal", legend.position="bottom")+
  coord_flip()

#ggplot(sum_sites, aes(x=count_taxa, y=mean_cover))+
#  geom_point(aes(color=condition, shape=site_name), size=3)+
#  scale_y_continuous(limits=c(0, 1), labels=scales::percent)+
#  scale_x_continuous(limits=c(0, 10))+
#  labs(x="Unique Species Found", y="Fe Floc Percent Cover", color="Condition", shape="Sites")

ggplot(sum_sites, aes(x=biodiversity, y=mean_cover))+
  geom_point(aes(color=condition, shape=condition), size=3)+
  geom_smooth(method="lm")+
  scale_y_continuous(labels=scales::percent, limits=c(0, 1))+
  labs(title="Species Richness as a function of Iron Floc Coverage", x="Species Richness", y="Fe Floc Percent Cover", color="Condition", shape="Condition")+
  theme(legend.direction="horizontal", legend.position="bottom")+
  coord_flip()
```
Data manipulation to get count data and join with water quality dataset
```{r}
#summarize taxa data to total number of each taxa by site and condition
taxa2 <- taxa %>% count(taxa, site_name, condition, transect_number)

#combine with water quality data
join <- water_quality1 %>%
+ full_join(y=taxa2, by = c("site_name", "transect_number"))

richness <- join %>%
  group_by(site_name, transect_number, percent_cover, condition.x) %>%
  summarise(richness = n_distinct(taxa))%>%
  ungroup()
```

Analysis
```{r}
#richness
mod2 <- lm(richness ~ percent_cover*condition.x, data = richness)
anova(mod2)

#abundance
mod3 <- lm(n ~ percent_cover*condition.x*taxa, data = join)
anova(mod3)
  
```



## Question 3

Does the effect of iron flocculate on the abundance and diversity of macroinvertebrates depend on if a stream has been restored?

Plot the biodiversity, #of species, and #of individuals against site names, filled by condition

```{r}
ggplot(sum_sites, aes(x=factor(site_name), y=count_taxa, fill=factor(condition)))+
  geom_bar(stat="identity")+
  scale_x_discrete(limits=c("j1", "j3", "cc", "ccr", "ar", "hb"))+
  labs(title="Richness per Site", x="Site Names", y="Species Richness", fill="Condition")+
  theme(legend.direction="horizontal", legend.position="bottom")

ggplot(sum_sites, aes(x=factor(site_name), y=total, fill=factor(condition)))+
  geom_bar(stat="identity")+
  scale_x_discrete(limits=c("j1", "j3", "cc", "ccr", "ar", "hb"))+
  labs(title="Abundance per Site", x="Site Names", y="Abundance", fill="Condition")+
  theme(legend.direction="horizontal", legend.position="bottom")
```


## Data Along Streams

Using a scatterplot to show the water quality levels within a stream with the position on the X axis

```{r}
#ggplot(water_quality, aes(x=geo_order, y=temp, color=as.factor(analogue)))+
#         geom_point()+
#  geom_line(aes(linetype=site_name))
#  geom_text()

#ggplot(water_quality, aes(x=geo_order, y=d_o, color=analogue))+
#         geom_point()+
#  geom_line(aes(linetype=site_name))

#ggplot(water_quality, aes(x=geo_order, y=`d_o%`, color=analogue))+
#         geom_point()+
#  geom_line(aes(linetype=site_name))

#ggplot(water_quality, aes(x=geo_order, y=turbidity, color=as.character(site_name)))+
#         geom_point()+
#  geom_line(aes(linetype=as.character(analogue)))+
#  geom_vline(xintercept=c(3.5, 6.5))+
#  scale_x_continuous(breaks=NULL)+
#  labs(x="Sequential Data along Streams", y="Turbidity (cm)", color="site_name", linetype="analogue")+
#  scale_fill_manual(values=cbPalette)
## find a way to krump down on the top of the graph to cut off anything above 120cm

#ggplot(water_quality, aes(x=as.character(transect_number), y=percent_cover, color=as.character(site_name), na.rm=TRUE))+
#  geom_point()

# can I make it such that I have a geom_line(aes(y=percent_cover)) for each individual site name?

# change this so that color maps to the analogue, linetype to analogue, and *the lines are labled with the site names directly on the graph*
```


## Amy's code chunk

```{r}
barGraphStats <- function(data, variable, byFactorNames) {
  count <- length(byFactorNames)
  N <- aggregate(data[[variable]], data[byFactorNames], FUN=length)
  names(N)[1:count] <- byFactorNames
  names(N) <- sub("^x$", "N", names(N))
  mean <- aggregate(data[[variable]], data[byFactorNames], FUN=mean)
  names(mean)[1:count] <- byFactorNames
  names(mean) <- sub("^x$", "mean", names(mean))
  sd <- aggregate(data[[variable]], data[byFactorNames], FUN=sd)
  names(sd)[1:count] <- byFactorNames
  names(sd) <- sub("^x$", "sd", names(sd))
  preSummaryStats <- merge(N, mean, by=byFactorNames)
  finalSummaryStats <- merge(preSummaryStats, sd, by=byFactorNames)
  finalSummaryStats$se <- finalSummaryStats$sd / sqrt(finalSummaryStats$N)
  return(finalSummaryStats)
  
ggplot(data = barGraphStats(data = damage2, variable = "herbivore", byFactorNames = c("Area_type")), aes(x=Area_type, y=mean, fill=Area_type)) +
   geom_bar(stat='identity', position=position_dodge()) +
    geom_errorbar(aes(ymin=mean-se, ymax=mean+se), width = 0.2, position = position_dodge(0.9))
}
```



