---
title: "IF_analysis"
output:
  html_document:
    df_print: paged
date: "2024-07-26"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Adding packages

```{r}
library(readxl)
library(ggplot2)
library(tidyverse)
library(lubridate)
```


## Datasets

Reading in the data sets.

```{r}
taxa <- read_excel("~/Desktop/Iron Floc UA Project/iron_floc_impact_on_macroinvertebrate_biodiversity_taxa.xlsx")
taxa_sim <- read_excel("~/Desktop/Iron Floc UA Project/iron_floc_impact_on_macroinvertebrate_biodiversity_taxa.xlsx", 
    sheet = "simulated data DO NOT USE")
water_quality <- read_excel("~/Desktop/Iron Floc UA Project/iron_floc_impact_on_macroinvertebrate_biodiversity_water_quality.xlsx", 
    sheet = "data")
water_sim <- read_excel("~/Desktop/Iron Floc UA Project/iron_floc_impact_on_macroinvertebrate_biodiversity_water_quality.xlsx", 
    sheet = "simulate data DO NOT USE")
```


Taking a look at the formats of the data:

```{r}
str(taxa)
str(taxa_sim)
str(water_quality)

```



## Graphing

Creating graphs.

```{r}
ggplot(water_quality, aes(x=d_o, y=ferrous_iron))+
  geom_point(color="purple")+
  labs(x="Dissolved Oxygen", y= "Ferrous Iron", title="Dissolved Oxygen and Iron")+
  theme_classic()
```

## Bar Graph for iron by plot location

```{r}
ggplot(water_quality, aes(x=factor(plot_location), y=ferrous_iron, fill=factor(transect_number)))+
  geom_col(position="dodge")
```

## Question 1

Does stream restoration increase the presence of iron flocculate by creating low-flow, acidic, and high iron water conditions compared to unrestored streams?

Plot the ferrous iron, dissolved oxygen mg/l, dissolved oxygen percentage, conductivity, temperature, and turbidity by restored/unrestored. Dissolved oxygens can share a graph.
```{r}
ggplot(water_sim, aes(x=factor(condition), y=factor(ferrous_iron), fill=factor(site_name)))+
  geom_col(position="dodge")

ggplot(water_sim, aes(x=factor(analogue), y=factor(percent_cover), fill=factor(site_name)))+
  geom_col(position="dodge")
```

## Question 2

How does iron flocculate affect the diversity and abundance of macroinvertebrates?

Create a biodiversity object (number of species divided by number of individuals). Plot number of species and number of individuals by percent cover and ferrous iron. Four graphs total.

```{r}
AR <- taxa_sim %>% filter(site_name=="AR")
AR_diversity<-length(unique(AR$taxa))/nrow(AR)

HB <- taxa_sim %>% filter(site_name=="HB")
HB_diversity <- length(unique(HB$taxa))/nrow(HB)

CC <- taxa_sim %>% filter(site_name=="CC")
CC_diversity <- length(unique(CC$taxa))/nrow(CC)

J1 <- taxa_sim %>% filter(site_name=="J1")
J1_diversity <- length(unique(J1$taxa))/nrow(J1)

J3 <- taxa_sim %>% filter(site_name=="J3")
J3_diversity <- length(unique(J3$taxa))/nrow(J3)
```

## Question 3

Does the effect of iron flocculate on the abundance and diversity of macroinvertebrates depend on if a stream has been restored?

```{r}

```











