---
title: "IF_analysis"
output:
  html_document:
    df_print: paged
date: "2024-07-26"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#### IMPORTANT: FOR 120.1 TURBIDITY MAKE IT GO OFF-SCALE, AND MARK WITH A STAR

Ask Amy about this. There may also still be a problem with the sum_sites cbind, but it may? be fixed?? double check that

## Adding packages

```{r}
library(readxl)
library(ggplot2)
library(tidyverse)
library(lubridate)
library(RColorBrewer)
```


## Datasets, Palettes

Reading in the data sets, and set up colorblind-friendly color palettes

```{r}
taxa <- read_excel("~/Desktop/Iron Floc UA Project/iron_floc_impact_on_macroinvertebrate_biodiversity_taxa.xlsx")

water_quality <- read_excel("~/Desktop/Iron Floc UA Project/iron_floc_impact_on_macroinvertebrate_biodiversity_water_quality.xlsx", 
    sheet = "data")

cbPalette <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
```


## Taking a look at the formats of the data:

```{r}
str(taxa)
str(water_quality)

```


## Graphing

Creating graphs.

```{r}
ggplot(water_quality, aes(x=d_o, y=ferrous_iron))+
  geom_point(color="purple")+
  labs(x="Dissolved Oxygen", y= "Ferrous Iron", title="Dissolved Oxygen and Iron")+
  theme_classic()
```


## Dataframes and objects

Create the all_sites dataframe to summarize the site_name, analogue, number of species, number of individuals, biodiversity/species richness percent, and mean iron floc coverage percent.

```{r}
prcnt_cvr <- water_quality %>%
  group_by(site_name) %>%
  summarize(mean_cover=mean(as.numeric(percent_cover), na.rm = TRUE), flow=mean(as.numeric(flow_rate)))

##this created all_sites, if you need to summarize the total number of unique/total number of entries
all_sites <- taxa %>% group_by(site_name, analogue) %>%
  ##select(site_name, taxa) %>% not necesary, try if there are errors
  summarize(count_taxa=n_distinct(taxa), total=n()) %>%
  mutate(biodiversity=count_taxa/total)
##adds the biodiversity values

sum_sites <- merge(prcnt_cvr, all_sites)
```


## Question 1

Does stream restoration increase the presence of iron flocculate by creating low-flow, acidic, and high iron water conditions compared to unrestored streams?

Plot the ferrous iron, dissolved oxygen mg/l, dissolved oxygen percentage, conductivity, temperature, and turbidity by restored/unrestored. Dissolved oxygens can share a graph.

```{r}
ggplot(water_quality, aes(x=factor(analogue), y=factor(ferrous_iron), fill=factor(site_name)))+
  geom_col(position="dodge")

ggplot(water_quality, aes(x=factor(analogue), y=factor(percent_cover), fill=factor(site_name)))+
  geom_col(position="dodge")

ggplot(water_quality, aes(x=site_name, y=ferrous_iron))+
         geom_bar(stat="identity")

ggplot(water_quality, aes(x=site_name, y=d_o))+
         geom_bar(stat="identity")

## d_o% goes here, rename it to remove the %

ggplot(water_quality, aes(x=site_name, y=conductivity))+
         geom_bar(stat="identity")

ggplot(water_quality, aes(x=site_name, y=temp))+
         geom_bar(stat="identity")

ggplot(water_quality, aes(x=site_name, y=turbidity))+
         geom_bar(stat="identity")

ggplot(prcnt_cvr, aes(x=site_name, y=mean_cover))+
         geom_bar(stat="identity")

ggplot(water_quality, aes(x=site_name, y=flow_rate))+
         geom_bar(stat="identity")

ggplot(water_quality, aes(x=ferrous_iron, y=percent_cover))+
  geom_bar(stat="identity")
## remove na's, invert graph

ggplot(sum_sites, aes(x=as.numeric(flow), y=as.numeric(mean_cover)))+
  geom_point()
```


## Question 2

How does iron flocculate affect the diversity and abundance of macroinvertebrates?

Plot number of species and number of individuals by percent cover and ferrous iron. Four graphs total. A fifth graph can show the biodiversity number against percent cover/ferrous iron.

```{r}
##change the all_sites and prcnt_cvr to be just all_sites, specifically reqrite the prcnt_cvr create code so all_sites gets the piping instead of prcnt_cvr
```


## Question 3

Does the effect of iron flocculate on the abundance and diversity of macroinvertebrates depend on if a stream has been restored?

Plot the biodiversity, #of species, and #of individuals against site names, filled by condition

```{r}
ggplot(all_sites, aes(x=analogue, y=count_taxa, fill=reorder(factor(site_name), analogue)))+
  geom_col(position="dodge")+
  labs(x="Site Names", y="Unique Species", fill="Analogue")+
  scale_fill_manual(values=cbPalette)+
  expand_limits(x=c(1, 3))

ggplot(all_sites, aes(x=analogue, y=total, fill=reorder(factor(site_name), analogue)))+
  geom_col(position="dodge")+
  labs(x="Analogue", y="Individuals Found", fill="Site Names")+
  scale_fill_manual(values=cbPalette)+
  expand_limits(x=c(1, 3))

ggplot(all_sites, aes(x=analogue, y=biodiversity, fill=reorder(factor(site_name), analogue)))+
  geom_col(position="dodge")+
  labs(x="Analogue Group", y="Biodiversity (%)", fill="Site Name")+
  scale_fill_manual(values=cbPalette)+
  expand_limits(x=c(1, 3), y=c(0, 0.5))+
  scale_y_continuous(labels=scales::percent)
```


## Data Along Streams

Testing using a scatterplot to show the water quality levels within a stream with the position on the X axis

```{r}
ggplot(water_quality, aes(x=geo_order, y=temp, color=analogue))+
         geom_point()+
  geom_line(aes(linetype=site_name))+
  labs(x="Sequential Data along Streams", y="Temperature (C)", color="Site Names", linetype="Site Names")

ggplot(water_quality, aes(x=geo_order, y=ferrous_iron, color=analogue))+
         geom_point()+
  geom_line(aes(linetype=site_name))+
  labs(x="Sequential Data along Streams", y="Dissolved Ferrous Iron (ppm)", color="Site Names", linetype="Site Names")

ggplot(water_quality, aes(x=geo_order, y=d_o, color=analogue))+
         geom_point()+
  geom_line(aes(linetype=site_name))+
  labs(x="Sequential Data along Streams", y="Dissolved Oxygen (mg/L)", color="Site Names", linetype="Site Names")

ggplot(water_quality, aes(x=geo_order, y=`d_o%`, color=analogue))+
         geom_point()+
  geom_line(aes(linetype=site_name))+
  labs(x="Sequential Data along Streams", y="Dissolved Oxygen (%)", color="Site Names", linetype="Site Names")

ggplot(water_quality, aes(x=geo_order, y=conductivity, color=analogue))+
         geom_point()+
  geom_line(aes(linetype=site_name))+
  labs(x="Sequential Data along Streams", y="Conductivity (uS/cm)", color="Site Names", linetype="Site Names")

ggplot(water_quality, aes(x=geo_order, y=turbidity, color=as.character(site_name)))+
         geom_point()+
  geom_line(aes(linetype=as.character(analogue)))+
  labs(x="Sequential Data along Streams", y="Turbidity (cm)", color="site_name", linetype="analogue")+
  scale_fill_manual(values=cbPalette)
# find a way to krump down on the top of the graph to cut off anything above 120cm
# change this so that color maps to the analogue, linetype to analogue, and *the lines are labled with the site names directly on the graph*
```

## Scatterplots of WQ data plotted against eachother to identify trendlines

to make sure na.rm works, check that the applicable variable is being read as numeric (or maybe another type?)

```{r}
ggplot(water_quality, aes(x=d_o, y=`d_o%`))+
  geom_point()+
  geom_smooth(method="lm")

ggplot(water_quality, aes(x=ferrous_iron, y=as.numeric(percent_cover), na.rm=TRUE))+
geom_point()+
  geom_smooth(method="lm")+
  labs(x="Dissolved Ferrous Iron (ppm)", y="Iron Floc Percent Cover")+
  expand_limits(x=c(0, 15)) # why doesn't this work? do i need a whole new function, or do i need to add an as. clause?

ggplot(water_quality, aes(x=flow_rate, y=as.numeric(percent_cover), na.rm=TRUE))+
  geom_point()+
  geom_smooth(method="lm")
# change default colors to cbPalette?
```




